---
title: "Diabetic_Readmission"
author: "Tiffany Chua"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Working Directory}
getwd()
```

```{r Libraries}
library(Amelia)
library(ggplot2)
library("GGally")
library(caTools)
library(caret)
library(class)
library("e1071")
library(psych)
library(corrplot)
library("ISLR")
library("cluster")
library(datasets)
library("factoextra")
library(dplyr)
library(rpart)
library(pROC)
library("rpart.plot")
library("randomForest")
library(pastecs)
library(tidyr)
library(Hmisc)
library(tidyverse)
library(car)
library(varImp)
library(stats)
library(data.table)
library(DT)
```

# Diabetes Readmission Project
Dataset Information:
The dataset represents 10 years (1999-2008) of clinical care at 130 US hospitals and integrated delivery networks. It includes over 50 features representing patient and hospital outcomes. Information was extracted from the database for encounters that satisfied the following criteria.

* (1) It is an inpatient encounter (a hospital admission).
* (2) It is a diabetic encounter, that is, one during which any kind of diabetes was entered to the system as a diagnosis.
* (3) The length of stay was at least 1 day and at most 14 days.
* (4) Laboratory tests were performed during the encounter.
* (5) Medications were administered during the encounter.

The data contains such attributes as patient number, race, gender, age, admission type, time in hospital, medical specialty of admitting physician, number of lab test performed, HbA1c test result, diagnosis, number of medication, diabetic medications, number of outpatient, inpatient, and emergency visits in the year before the hospitalization, etc.

As the dataset has unusual characters (?) in place of NAs, I will change the string to NA upon loading the dataset.
```{r Load Dataset}
dataset <- read.csv("./diabetic_data.csv",
                    na.string = "?",
                    header = T)
head(dataset)
```

## ID Mapping

Categorical/factor variables:

* admission_type_id
  * 1. Emergency
  * 2. Urgent
  * 3. Elective
  * 4. Newborn
  * 5. Not available  --> NA
  * 6. NULL           --> NA
  * 7. Trauma Center
  * 8. Not Mapped     --> NA
  
* discharge_disposition_id
  * 1. Discharged to home
  * 2. Discharged/transferred to another short term hospital
  * 3. Discharged/transferred to SNF
  * 4. Discharged/transferred to ICF
  * 5. Discharged/transferred to another type of inpatient care institution
  * 6. Discharged/transferred to home with home health service
  * 7. Left AMA 
  * 8. Discharged/transferred to home under care of Home IV provider
  * 9. Admitted as an inpatient to this hospital
  * 10. Neonate discharged to another hospital for neonatal aftercare
  * 11. Expired
  * 12. Still patient or expected to return for outpatient services
  * 13. Hospice/Home
  * 14. Hospice/Medical facility
  * 15. Discharged/transferred within this institution to Medicare approved swing bed 
  * 16. Discharged/transferred/referred another institution for outpatient services
  * 17. Discharged/transferred/referred to this institution for outpatient services
  * 18. NULL 
  * 19. Expired at home. Medicaid only, hospice
  * 20. Expired in a medical facility. Medicaid only, hospice.
  * 21. Expired, place unknown. Medicaid only, hospice.
  * 22. Discharged/transferred to another rehab fac including rehab units of a hospital .
  * 23. Discharged/transferred to a long term care hospital.
  * 24. Discharged/transferred to a nursing facility certified under Medicaid but not certified under Medicare.
  * 25. Not Mapped
  * 26. Unknown/Invalid
  * 27. Discharged/transferred to a federal health care facility.
  * 28. Discharged/transferred/referred to a psychiatric hospital of psychiatric distinct part unit of a hospital
  * 29. Discharged/transferred to a Critical Access Hospital (CAH).
  * 30. Discharged/transferred to another Type of Health Care Institution not Defined Elsewhere
  
* admission_source_id
  * 1. Physician Referral
  * 2. Clinic Referral
  * 3. HMO Referral
  * 4. Transfer from a hospital
  * 5. Transfer from a Skilled Nursing Facility (SNF)
  * 6. Transfer from another health care facility
  * 7. Emergency Room
  * 8. Court/Law Enforcement
  * 9. Not Available
  * 10. Transfer from critial access hospital
  * 11. Normal Delivery
  * 12. Premature Delivery
  * 13. Sick Baby
  * 14. Extramural Birth
  * 15. Not Available
  * 16. 
  * 17. NULL
  * 18. Transfer From Another Home Health Agency
  * 19. Readmission to Same Home Health Agency
  * 20. Not Mapped
  * 21. Unknown/Invalid
  * 22. Transfer from hospital inpt/same fac reslt in a sep claim
  * 23. Born inside this hospital
  * 24. Born outside this hospital
  * 25. Transfer from Ambulatory Surgery Center
  * 26. Transfer from Hospice

Other column notes:

* Columns to remove:
  * $encounter_id
  * $patient_nbr
  * $payer_code
  * $medical_specialty (maybe?)

